---
title: "Tidy Tuesday"
author: "Annie Deck"
date: "2024-12-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      warning=FALSE)
```
# U.S. Customs and Border Protection Encounter Data

* This week for Tidy Tuesday we are working with the U.S. Customs and Border Protection (CBP) Encounter Data 
* I wanted to learn some new tools to make the aesthetics of a figure look really nice. 
* Since the numbers on the y-axis were very large and hard to interpret quickly, I learned how to insert breaks and custom labels. I also used a new ggplot theme to only retain the lines on the x and y axes and then I added very light horizontal grey lines to make the figure easy to interpret.



# Load Libraries
```{r}
library(here)
library(tidyverse)
library(viridis) 
library(usmap)
```

# Bring in data
```{r}
##read in from Github
cbp_resp <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-11-26/cbp_resp.csv')

cbp_state <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-11-26/cbp_state.csv')


## save to data folder 
readr::write_csv(cbp_resp, here::here("Tidy_Tuesday", "Customs", "Data", "cbp_resp.csv"))

readr::write_csv(cbp_state, here::here("Tidy_Tuesday", "Customs", "Data", "cbp_state.csv"))

```
# Data prep
```{r}
cbp_resp_sumarized <- cbp_resp %>%
  group_by(fiscal_year, encounter_type) %>% ##group by year and encounter type
  summarise(total_encounters = sum(encounter_count, na.rm = TRUE)) %>% ##make a new column with the sum of each enounter type for each year
  ungroup()

```


# Plotting 
```{r, fig.align='center'}
encounters_plot <- cbp_resp_sumarized %>%
  ggplot(aes(x= fiscal_year,
             y= total_encounters,
             color= encounter_type)) +
  geom_line() +
  geom_point(size=3.3)+
  scale_y_continuous(
    breaks = c(500000, 1000000, 1500000), ##edit y axis to have breaks at these points
    labels = c("500k", "1M", "1.5M")) + ##custom y axis lables 
  scale_color_manual(values = c("#FFB6C1","#AFEEEE", "#2166ac")) + ##make lines & points nice colors
  theme_classic() + ##add theme that takes away a lot of the unnecessary grid lines
  theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.5), ##make title look nice
        plot.subtitle = element_text(size= 14, hjust=0.5), #make subtitle appropiate size and center it 
        legend.position = "top", ##move legend to the top
        legend.title=element_blank(), #remove unnecessary legend title
        panel.grid.major.y = element_line(color = "lightgrey", ##add thin horizontal grey lines
                                           linewidth = 0.2),
        axis.text.x = element_text(size = 11),  ##make axis lables larger and easier to read
        axis.text.y = element_text(size = 11)) +  
  labs(x = "", ##axis lables seem redundant so removing these
       y = "",
       title = "U.S. Border Analysis (FY 2020-2024)",
       subtitle = "Trends in Total Encounters",
       caption = "Source: Tidy Tuesday's U.S. Customs and Border Protection (CBP) Encounter Data") 

encounters_plot

```

# Saving figure
```{r}
ggsave(here("Tidy_Tuesday", "Customs", "Output", "encounters_plot.png"), width= 8, height=5) ##save to Output folder. edits dimensions to make it long and skinny

```

